---
title: "Creating figures for the paper"
output: 
  github_document:
    toc: true
    toc_depth: 5
date: "2025-11-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This R markdown shows how all figure panels in the paper were made, except for Figure 2, which was created in PowerPoint. Before running this code, be sure to unzip all compressed folders in `data/`

`multilocus-suppression-functions.R` contains all functions used for these plots. See the each function's specs for a detailed breakdown of how it works.

Panels were combined using Adobe Illustrator 2025.

## Single locus plots

* Note: when the mean fitness of the population was reduced to 0 (such that the genetic load imposed by the suppression drive was 1), the WF SLiM simulation crashed because 10,000 offspring could not be produced. For such replicates, we manually ensured that csvs summarizing the replicates had an ending mean fitness of 0 and genetic load of 1.

```{r eval=F}
rm(list = ls())
if (!requireNamespace("here", quietly = TRUE)) install.packages("here")
library(here)

source(here("multilocus-suppression-functions.R"))

# Calculate necessary text size, based on the size of the Adobe Illustrator artboard 
# and the amount of scaling done in Overleaf
sizes = calculate_text_size(1080, 1, 11)
text_size = sizes$text_size

max_gen = 100
linewidth = 1

# dsx and s_d = 0.2 for resistance rate of 1e-5

# legend.res = plot_single_locus_joint_gl_joint_res(resistance_rate = 1e-5,
#                                             max_gen = max_gen,
#                                             legend.pos = "right",
#                                             text_size = text_size,
#                                             linewidth = linewidth)
# ggsave(plot = legend.res$gl, filename = "legend-single-locus.pdf", width = 6, height = 6)

plts.res1e_5 = plot_single_locus_joint_gl_joint_res(resistance_rate = 1e-5,
                                            max_gen = max_gen,
                                            legend.pos = "none",
                                            text_size = text_size,
                                            linewidth = linewidth)


panelA = plts.res1e_5$gl + xlab("") + ggtitle(TeX(r"($r_d = 10^{-5}$)")) + 
  theme(plot.title = element_text(size = text_size, hjust = 0.5))

panelB = plts.res1e_5$res

# dsx and s_d = 0.2 for resistance rate of 1e-4
plts.res1e_4 = plot_single_locus_joint_gl_joint_res(resistance_rate = 1e-4,
                                            max_gen = max_gen,
                                            legend.pos = "none",
                                            text_size = text_size,
                                            linewidth = linewidth)

panelC = plts.res1e_4$gl + xlab("") + ylab("") + ggtitle(TeX(r"($r_d = 10^{-4}$)")) + 
  theme(plot.title = element_text(size = text_size, hjust = 0.5)) 

panelD = plts.res1e_4$res + ylab("")

# dsx and s_d = 0.2 for resistance rate of 1e-3
plts.res1e_3 = plot_single_locus_joint_gl_joint_res(resistance_rate = 1e-3,
                                            max_gen = max_gen,
                                            legend.pos = "none",
                                            text_size = text_size,
                                            linewidth = linewidth)
panelE = plts.res1e_3$gl + xlab("") + ylab("") + ggtitle(TeX(r"($r_d = 10^{-3}$)")) + 
  theme(plot.title = element_text(size = text_size, hjust = 0.5)) 

panelF = plts.res1e_3$res + ylab("")


# combine panels
dist.w = -0.18
dist.h = -0.1

combined = plot_grid(panelA, NULL, panelC, NULL, panelE,
                     NULL, NULL, NULL, NULL, NULL,
                     panelB, NULL, panelD, NULL, panelF,
                     nrow = 3, ncol = 5,
                     align = "hv",
                     rel_widths = c(1, dist.w, 1, dist.w, 1),
                     rel_heights = c(1, dist.h, 1))

# To save, uncomment the line below:

# ggsave(plot = combined,filename = "combined-gl-res-panel.pdf", width = 15, height = 8.75)
```

## Multilocus plot with unlinked targets

```{r eval=F, message = F}
rm(list = ls())
if (!requireNamespace("here", quietly = TRUE)) install.packages("here")
library(here)

source(here("multilocus-suppression-functions.R"))

# Calculate necessary text size, based on the size of the Adobe Illustrator artboard 
# and the amount of scaling done in Overleaf
sizes = calculate_text_size(900, 1, 11)
text_size = sizes$text_size

# Load in summary of the cluster run varying n and lambda^*
# lambda^* is indicated by the FGL column ("fixed genetic load")
combined_data = read_csv(here("data/n-lambda/cas9-drive1-target0.8/n-vs-lambda-drive1-targets0.8-summary.csv"),
                         col_names = T)


# Panels A and B: genetic load and drive and target frequencies 
max_gen = 100; linewidth = 1
palette = c("saddlebrown", "green4", "mediumslateblue") # 1 5 10

title = TeX(r"($\lambda^* = 0.95$)")

res.no.legend = plot_multilocus_genetic_load_summaries(text_size = text_size, max_gen = max_gen,
                                             linewidth = linewidth, title = "", legend.pos = "none")

panelA = res.no.legend$gl.plot 
panelB = res.no.legend$drive.and.target.plot

# Panel C: heatmap of max genetic load
point.size = 3

x_tick_marks = seq(1, 10, by = 2)
y_tick_marks = seq(0.5, 0.95, by = 0.1)
xindex = 2; yindex = 1
z_tick_marks = y_tick_marks

mingl = min(combined_data$avg_max_load) # 0.459
maxgl = max(combined_data$avg_max_load) # 0.94

zmin = 0.45; zmax = 0.95
heat.maxGL = heatmap_detailed(csv = combined_data, xindex = xindex, yindex = yindex, zindex = 15,
                       x_tick_marks = x_tick_marks, y_tick_marks = y_tick_marks,
                       z_tick_marks = z_tick_marks, zlow = zmin, zhigh = zmax,
                       color_scale = c("white", "red"),  xl = TeX(r"(Number of target sites ($n$))"), 
                       yl = TeX(r"(Potential genetic load ($\lambda^*$))"),
                       zl = TeX(r"($\max(\lambda(t))$)"),
                       title = "", text_size = text_size, legend.pos = "right") +
  geom_point(aes(x = 1, y = 0.95), color = palette[1], size = point.size) +
  geom_point(aes(x = 5, y = 0.95), color = palette[2], size = point.size)+
  geom_point(aes(x = 10, y = 0.95), color = palette[3], size = point.size)

#ggsave(plot = heat.maxGL, filename = paste0("heatmap-max-load-legend.pdf"),
#       width = 6, height = 6)

panelC = heat.maxGL + theme(legend.position = "none")


# Panel D: heatmap of time until genetic load falls below 20%
zlow = floor(min(combined_data$avg_time_to_genetic_load_below_marker)) # 15
zhigh = ceiling(max(combined_data$avg_time_to_genetic_load_below_marker)) # 109
z_tick_marks = seq(15, 109, by = 30)

heat.persist = heatmap_detailed(csv = combined_data, xindex = xindex, yindex = yindex, zindex = 19,
                       x_tick_marks = x_tick_marks, y_tick_marks = y_tick_marks,
                       z_tick_marks = z_tick_marks, zlow = zlow, zhigh = zhigh,
                       color_scale = c("white", "dodgerblue"), xl = TeX(r"(Number of target sites ($n$))"), 
                       yl = TeX(r"(Potential genetic load ($\lambda^*$))"),
                       zl = TeX(r"($\tau$)"),
                       title = "", text_size = text_size, legend.pos = "right")+
  geom_point(aes(x = 1, y = 0.95), color = palette[1], size = point.size) +
  geom_point(aes(x = 5, y = 0.95), color = palette[2], size = point.size)+
  geom_point(aes(x = 10, y = 0.95), color = palette[3], size = point.size)

#ggsave(plot = heat.persist, filename = paste0("heatmap-time-to-below20-legend.pdf"),
#       width = 6, height = 6)

panelD = heat.persist + theme(legend.position = "none")


# Combine the panels
dist.h = -0.25
dist.w = 0.45

combined = plot_grid(panelA, NULL, panelC,
                     NULL, NULL, NULL,
                     panelB, NULL, panelD,
                     align = "hv", nrow = 3, ncol = 3,
                     rel_widths = c(1, dist.w, 1),
                     rel_heights = c(1, dist.h, 1))

# Save by uncommenting line below:

#ggsave(plot = combined, filename = paste0("combined-panel-n-vs-lambda-drive1-target0.8.pdf"), width = 12, height = 12)
```


## Multilocus plot with varying linkage (log jam plot)

Panels A-D:
```{r eval=F}
rm(list = ls())
if (!requireNamespace("here", quietly = TRUE)) install.packages("here")
library(here)

source(here("multilocus-suppression-functions.R"))

# Calculate necessary text size, based on the size of the Adobe Illustrator artboard 
# and the amount of scaling done in Overleaf
sizes = calculate_text_size(1080, 1, 11)
text_size = sizes$text_size

# Color palette for rho
rho.cols = rev(c(
  "black",
  "#440154FF",  
  "#414487FF",  
  "#2A788EFF", 
  "#22A884FF",  
  "#7AD151FF", 
  "#E5D510FF"   
))

max_gen = 100
linewidth = 1

dir = here("data/log-jam/lambda0.95-drive1-target0.8/")
data = read_csv(paste0(dir, "n-vs-rho-lambda0.95-drive1-targets0.8-summary.csv"), col_names = T)

# Get plots summarizing the trends
res.jam =  log_jam_summary_plots(data = data,
                                 text_size = text_size, linewidth = 1, 
                                 point_size = 2.5, rho_colors = rho.cols)
#ggsave(plot= res.jam$max.gl, filename = "legend-log-jam.pdf", width = 6, height = 6)


# Panels A and B: max genetic load and time until this state
panelA = res.jam$max.gl + theme(legend.position = "none") +
  scale_y_continuous(breaks = seq(0.5, 1, length.out = 6), limits = c(0.5,1))

panelB = res.jam$time.max.gl + theme(legend.position = "none") +
  scale_y_continuous(breaks = seq(10,15, by = 1), limits = c(10,15.5))

# Panels C and D: ending genetic load and avg time to GL < 20% (only showing params where this happened more than once)
panelC = res.jam$ending.gl + theme(legend.position = "none")

panelD = res.jam$time.gl.below.0.2.given.rate.above.10 + theme(legend.position = "none") +
  scale_y_continuous(breaks = seq(0, 6000, by = 2000), limits = c(0, 6300))


# Combine plots
dist.w = 0.05; dist.h = 0
combined = plot_grid(panelA, NULL, panelB,
                     NULL, NULL, NULL,
                     panelC, NULL, panelD,
                     nrow = 3, ncol = 3, 
                     align = "hv", rel_widths = c(1, dist.w, 1),
                     rel_heights = c(1, dist.h, 1))

# Save by uncommenting line below:
ggsave(plot = combined, width = 14.73462, height = 14, filename = paste0("log-jam-panels-ABCD.pdf"))
```

Panels E and F: replicates of n = 4, rho = 1e-7. Note: I added haplotype diagrams in PowerPoint.

```{r eval=F}
rm(list = ls())
if (!requireNamespace("here", quietly = TRUE)) install.packages("here")
library(here)

source(here("multilocus-suppression-functions.R"))

# Calculate necessary text size, based on the size of the Adobe Illustrator artboard 
# and the amount of scaling done in Overleaf
sizes = calculate_text_size(1080, 1, 11)
text_size = sizes$text_size

base = here("data/log-jam/lambda0.95-drive1-target0.8/") # where data is stored

# Panel E: log jam but escapes
data4.1 = read_csv(here(paste0(base, "log-jam-1e-7/num_targets4/","20251101122549_6412581988158431629_output.csv")))
p4.1 = plot_multilocus_single_replicate(data4.1, 
                                        add_genetic_load = F, 
                                        add_frequencies = T,
                                        text_size = text_size,
                                        linewidth = 1, alpha = 1)

# Save by uncommenting line below:
# ggsave(plot = p4.1, filename = paste0("raw-legend-log-jam-4-targets-rho1e-7-rep1.pdf"), width = 8, height = 8)

# Panel F: perpetual log jam
data5 = read_csv(here(paste0(base, "log-jam-1e-7/num_targets4/","20251101142128_5898902782196442742_output.csv")))
p4.2 = plot_multilocus_single_replicate(data5, 
                                        add_genetic_load = F, 
                                        add_frequencies = T,
                                        text_size = text_size,
                                        linewidth = 1, alpha = 1)

# Save by uncommenting line below:

# ggsave(plot = p4.2 + scale_x_continuous(expand = expansion(mult = 0.1)), filename = paste0("raw-log-jam-4-targets-rho1e-7-rep2.pdf"),width = 8, height = 8)
```




